Lies and the lying liars who tell them

### 2018-11-03 GSV-comment-brief.Rmd
Use checkpoint; other textual tweaks

---
title: "True lies"
subtitle: "Comment on Garbarino, Slonim and Villeval (2018)"
author: "David Hugh-Jones"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{setspace}
  - \usepackage{placeins}
  - \onehalfspacing
  - \usepackage{fontspec}
output:
  pdf_document:
    md_extensions: +raw_attribute
    latex_engine: xelatex
  tufte::tufte_handout:
    latex_engine: xelatex
    md_extensions: +raw_attribute
  bookdown::tufte_handout2:
    latex_engine: xelatex
    md_extensions: +raw_attribute
  bookdown::tufte_html2: 
    pandoc_args: ['--self-contained']
    keep_intermediates: true
bibliography: bibliography.bib
link-citations: yes
fontsize: 12pt
mainfont: Times New Roman
mathfont: Times New Roman
---

>  Garbarino, Slonim and Villeval (2018) describe a new method to calculate the probability distribution 
>  of the number of lies told in "coin flip" style experiments. I show that their estimates and
>  confidence intervals are flawed. I explain why, and describe two better
>  ways to estimate the probability distribution of the proportion of liars.

@@

@@

# set to FALSE to recalculate all confidence interval estimates; TRUE for speed:
cache_results <- TRUE

@@

First, if 10 out of 10 heads are reported, you should learn two things. On the one hand, there are
probably many liars in your sample. On the other hand, probably a lot of coins really landed heads.
The probability distribution in (\ref{eqn:truncated-binom}) does not take account of this. In
particular, for *R = N*, it reduces to the binomial distribution. This is not correct. For example,
in my simulations with $P = 0.5$, the probability that there were 10 heads when $R = N = 10$ was
about 1 in `r inv_prob_h`, not 1 in 1024. When $\\lambda = 0.2$, it is about 1 in 
`r inv_prob_h_l20`. The same is true at the other end of the scale: when *R* is low, we indeed learn
that $T \le R$ but also that the overall distribution is not shaped like the prior.

@@

The first of these two problems means that the GSV distribution of *Lies* is wrong. In the appendix
I show that their estimator can have substantial bias. In fact, it performs worse than the naïve
estimator from (\ref{eqn:naive-estimator}), $\frac{R-(1-P)}{P}$. Also, the GSV confidence intervals do not always achieve nominal coverage of *Lies*. When the number of heads reported is either high or low, the percentage of confidence intervals containing *Lies* may fall below the nominal value.

The second problem means that the GSV confidence intervals are not correct for $\lambda$. We can
tell that the second problem matters, because GSV coverage of $\lambda$ is much worse than coverage 
of *Lies*. Also, since the GSV estimator of *Lies* is biased, it is a biased estimator of
$\lambda$.

```{r disagg-coverage}
disagg <- sim_est %>% 
      filter(method=="gsv") %>% 
      group_by(
        `R/N:` = R/N, N, P, CI
      ) %>% 
      summarize(mean_in_ci = mean(within_ci, na.rm = TRUE), prop = n()/nrow(.)) %>% 
      mutate(CI = paste0("CI ", CI, "%")) %>% 
      spread(CI, mean_in_ci) %>% 
      mutate(ok  = `CI 95%` > 0.95)

@@

The problem is especially bad when there are many reports of heads. In this case there were probably
many true heads, so *T* is high and the true sample size *N - T* is low. Table \ref{tab:coverage-rn}
shows this. It splits the simulations by the proportion of reported heads, *R/N*. Coverage levels
fall off sharply as *R/N* increases. Note that for fair coin flips, *R/N* is typically greater than
0.5, in the simulations and in reality.

@@

as_huxtable(coverage_N, add_colnames = TRUE) %>% 
      set_rowspan(c(2, 4, 6), "Method", 2) %>% 
      set_valign(c(2, 4, 6), "Method", "top") %>% 
      set_number_format(-1, contains("CI"), list(pct_format)) %>% 
      set_caption("Confidence interval coverage by sample size") %>% 
      set_label("tab:coverage-N") %>% 
      theme_plain() %>% 
      set_background_color(everywhere, 1, NA) %>% 
      set_top_border(c(4, 6), everywhere, 0.4) %>% 
      set_bottom_border(6, 1, 0.4)
```

@@

sim_est %>% 
      filter(CI == 90) %>% 
      group_by(method, N) %>% 
      summarize(mean_sq_error = mean(error_sq, na.rm = TRUE)) %>% 
      spread(N, mean_sq_error, sep = ": ") %>% 
      ungroup() %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      as_huxtable(add_colnames = TRUE) %>% 
      set_caption("Mean squared errors by method and N") %>% 
      set_label("tab:point-estimates") %>% 
      theme_plain()
```

@@

nat_means <- sapply(distribs, dist_mean)
nat_cis <- sapply(distribs, dist_quantile, probs = c(0.025, 0.975))
rownames(nat_cis) <- c("ci.low", "ci.high")

@@

ggplot(nats, aes(x = factor(resnat, levels = resnat), y = mean, fill = mean)) +
      geom_col(alpha = 0.8) +
      geom_linerange(aes(ymin = ci.low,ymax = ci.high)) +
      geom_point(aes(y = prop_liars), shape = 4, size = 3, colour = "red") +
      xlab("Nation") + ylab("Prop. liars") +
      viridis::scale_fill_viridis(option = "E", guide = FALSE) + 
      coord_flip()

>>>>

---
title: "Lies and the lying liars who tell them"
subtitle: "Comment on Garbarino, Slonim and Villeval (2018)"
author: "David Hugh-Jones"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{setspace}
  - \usepackage{placeins}
  - \onehalfspacing
  - \usepackage{fontspec}
output:
  pdf_document:
    md_extensions: +raw_attribute
    latex_engine: xelatex
  tufte::tufte_handout:
    latex_engine: xelatex
    md_extensions: +raw_attribute
  bookdown::tufte_handout2:
    latex_engine: xelatex
    md_extensions: +raw_attribute
  bookdown::tufte_html2: 
    pandoc_args: ['--self-contained']
    keep_intermediates: true
bibliography: bibliography.bib
link-citations: yes
fontsize: 12pt
mainfont: Times New Roman
mathfont: Times New Roman
---

> Garbarino, Slonim and Villeval (2018) describe a new method to calculate the probability
> distribution of the proportion of lies told in "coin flip" style experiments. I show that their
> estimates and confidence intervals are flawed. I demonstrate two better ways to estimate the 
> probability distribution of what we are really interested in: the proportion of liars.

@@

library(checkpoint)
checkpoint("2018-11-03")

@@

# set to FALSE to recalculate all confidence interval estimates; TRUE for speed:
cache_results <- FALSE

@@

First, if 10 out of 10 heads are reported, you should learn two things. On the one hand, there are
probably many liars in your sample. On the other hand, probably a lot of coins really landed heads.
The probability distribution in equation (\ref{eqn:truncated-binom}) does not take account of this.
In particular, for *R = N*, it reduces to the binomial distribution. This is not correct. For
example, in my simulations with $P = 0.5$, the overall probability that there were 10 heads when $R
= N = 10$ was about 1 in `r inv_prob_h`, not 1 in 1024. When $\lambda = 0.2$, it is about 1 in 
`r inv_prob_h_l20`. The same is true at the other end of the scale: when *R* is low, we indeed learn
that $T \le R$ but also that the overall distribution is not shaped like the prior.

@@

The first of these two problems means that the GSV distribution of *Lies* is wrong. In the appendix
I show that their estimator can have substantial bias. In fact, it performs worse than the naïve
estimator from equation (\ref{eqn:naive-estimator}), $\frac{R-(1-P)}{P}$. Also, the GSV confidence
intervals do not always achieve nominal coverage of *Lies*. When the number of heads reported is
either high or low, the percentage of confidence intervals containing *Lies* may fall below the
nominal value.

The second problem means that the GSV confidence intervals are not correct for $\lambda$. We can
tell that the second problem matters, because the confidence interval coverage of $\lambda$ is much
worse than coverage of *Lies*. The GSV estimator is also a biased estimator of $\lambda$.

```{r disagg-coverage}

disagg <- sim_est %>% 
      filter(method=="gsv") %>% 
      group_by(
        `R/N:` = R/N, N, P, CI
      ) %>% 
      summarize(mean_in_ci = mean(within_ci, na.rm = TRUE), prop = n()/nrow(.)) %>% 
      mutate(CI = paste0("CI ", CI, "%")) %>% 
      spread(CI, mean_in_ci) %>% 
      mutate(ok  = `CI 95%` > 0.95)

@@

The problem is especially bad when there are many reports of heads. In this case there were probably
many true heads, so *T* is high and the true sample size $N - T$ is low. Table \ref{tab:coverage-rn}
shows this. It splits the simulations by the proportion of reported heads, *R/N*. Coverage levels
fall off sharply as *R/N* increases. Note that for fair coin flips, *R/N* is typically greater than
0.5, in the simulations and in reality.

@@

as_huxtable(coverage_N, add_colnames = TRUE) %>% 
      set_rowspan(c(2, 4, 6), "Method", 2) %>% 
      set_valign(c(2, 4, 6), "Method", "top") %>% 
      set_number_format(-1, contains("CI"), list(pct_format)) %>% 
      set_caption("Confidence interval coverage by sample size") %>% 
      set_label("tab:coverage-N") %>% 
      theme_plain() %>% 
      set_background_color(everywhere, 1, NA) %>% 
      set_top_border(c(4, 6), everywhere, 0.4) %>% 
      set_bottom_border(6, 1, 0.4) %>% 
      set_right_border(everywhere, 1, 0.4)
```

@@

sim_est %>% 
      filter(CI == 90) %>% 
      mutate(
        Method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian"),
        method = NULL
      ) %>% 
      group_by(Method, N) %>% 
      summarize(mean_sq_error = mean(error_sq, na.rm = TRUE)) %>% 
      spread(N, mean_sq_error, sep = ": ") %>% 
      ungroup() %>% 
      as_huxtable(add_colnames = TRUE) %>% 
      set_caption("Mean squared errors by method and $N$") %>% 
      set_label("tab:point-estimates") %>% 
      theme_plain()
```

@@

nat_means <- sapply(distribs, dist_mean)
nat_cis <- sapply(distribs, dist_hdr, conf_level = 0.95)
rownames(nat_cis) <- c("ci.low", "ci.high")

@@

ggplot(nats, aes(x = factor(resnat, levels = resnat), y = mean, fill = mean)) +
      geom_col(alpha = 0.8) +
      geom_linerange(aes(ymin = ci.low,ymax = ci.high)) +
      geom_point(aes(y = prop_liars), shape = 4, size = 3, colour = "red") +
      xlab("Nation") + ylab("Prop. liars") +
      scale_fill_viridis_c(option = "A", guide = "none") + 
      theme_light()+
      coord_flip()
